% Hamad Medical Corporation
% Georges Younes

\setcounter{chapter}{-1}
\chapter{Summary}\label{chp:summary}

\section{Objective(s)}
Maximum of 750 characters. Introduce the subject and motivation of the research to the reader in such a way that it can be reasonably understood by the lay person. Define the problem that the research has addressed, and provide a statement describing the impetus for the research.
\\\hrule

\section{Justification(s)}
Maximum of 750 characters. Describe the justifications for conducting the research.
\\\hrule

\section{Methodology}
Maximum of 1,250 characters. State the approach taken by the research Team to solve the problem(s). The statement can make mention where applicable, of the use experimental design and procedures, design of surveys, and/or any simulation conducted.
\\\hrule

\section{Conclusions of Results}
Maximum of 2,000 characters. This section is intended to present the most significant findings of the research project, and to outline their wider implications and draw conclusions.
\\\hrule

\section{Significance}
Maximum of 750 characters. State the most significant contribution(s) of the research to the discipline.
\\\hrule

\section{Benefits to Qatar}
Maximum of 750 characters. Make reference to project alignment with Qatar National Research Strategy (QNRS) and how and in what way the state of Qatar can benefit from the research outcomes.
\\\hrule

\clearpage

\section{Summary Slide}

\subsection{Major Impacts}
\begin{enumerate}
  \item Impact 1
  \item Impact 2
\end{enumerate}

\subsection{Project/Technology Description}
\subsubsection{Major Challenges Addressed}
\subsubsection{Approach/Solution}

\subsection{Technical Approach and Solution}
Briefly describe the technology. May use graphic illustrations.

\subsubsection{Objective}
\subsubsection{Key Results}
\subsubsection{Intellectual Property Position}

\subsection{Benefits/Potential Applications/Customers/Markets}
\subsubsection{Benefits}
\subsubsection{Potential Applications}
\subsubsection{Customers/Market}

\subsection{Key Outcomes}
\begin{enumerate}
  \item Impact 1
  \item Impact 2
\end{enumerate}
